Temporal graph realization from fastest paths

### 2024-01-19 SAND2024/SAND24-shortversion.tex
Removed folder for figures from SAND
SAND submission - first version

\clearpage

@@

We obtain the following results. We show that \deltaExact\ is W[1]-hard when parameterized by the \emph{feedback vertex number} of the underlying graph (\cref{thm:W1wrtFVS}). 
To this end, we first give a reduction from \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval} to a variant of \deltaExact\ where the period $\Delta$ is infinite, that is, when the labeling is non-periodic. We use a special gadget (the ``infinity'' gadget) which allows us to transfer the result to a finite period $\Delta$. The latter construction is independent from the particular reduction we use, and can hence be treated as a reduction from the non-periodic to the periodic setting.
Note that our parameterized hardness result with respect to the feedback vertex number 
also implies W[1]-hardness for any smaller parameter, such as \emph{treewidth}, \emph{degeneracy}, \emph{cliquewidth}, 
\emph{distance to chordal graphs}, and \emph{distance to outerplanar graphs}.

@@

There are some problem settings that share similarities with ours, which we discuss now in more detail.

@@

\begin{theorem}[$\star$]\label{thm:W1wrtFVS}
    \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$ (if not we can add $k \max_{i,j}|F_{i,j}|$ vertices to each $W_i$ and use those to add up to $\max_{i,j}|F_{i,j}|$ additional isolated edges to each $F_{i,j}$).
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

We give a reduction to a variant of \deltaExact\ where the period $\Delta$ is infinite (that is, the sought temporal graph is not periodic) and we allow $D$ to have infinity entries, meaning that the two respective vertices are not temporally connected. 
Note that, given the matrix $D$, we can easily compute the underlying graph $G$, as follows. Two vertices $v,v'$ are adjacent if $G$ if and only if $D_{v,v'}=1$, as having an edge between $v$ and $v'$ is the only way that there exists a temporal path from $v$ to $v'$ with duration 1. 
For simplicity of the presentation of the reduction, we describe the underlying graph $G$ (which directly implies the entries of $D$ where $D(v,v')=1$) and then we provide the remaining entries of $D$. 
At the end of the proof we show how to obtain the result for a finite $\Delta$ and a matrix $D$ of durations of fastest paths, that only has finite entries.

@@

Every \emph{edge selection gadget} is associated with a color combination $i,j$ in the \textsc{Multicolored Clique} instance, and its main purpose is to ``select'' an edge connecting a vertex from color $i$ with a vertex from color $j$.
Roughly speaking, the edge selection gadget consists of $m$ paths, one for every edge in $F_{i,j}$ (see \cref{fig:hardness1} for reference). The distance matrix $D$ will enforce that the labels on those paths effectively order them temporally, that is, in particular, the labels on one of the paths will be smaller than the labels on all other paths. The edge corresponding to this path is selected.

 We have a \emph{verification gadget} for every color $i$. They interact with the edge selection gadgets as follows.
 The verification gadget for color $i$ is connected to all edge selection gadgets that involve color $i$. More specifically, this is connected to every path corresponding to an edge at a position in the path that encodes the endpoint of color $i$ of that edge (again, see \cref{fig:hardness1} for reference). Intuitively, the distances in the verification gadget are only realizable if the selected edges all have the same endpoint of color $i$.
Hence, the distances of all verification gadgets can be realized if and only if the selected edges form a clique.

@@

    \emph{Connector Gadget.} Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by fast paths, which will make arguing about possible labelings in \textsc{Yes}-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and two connector gadgets for each validation gadget.

    The \emph{first connector gadget for the edge selection gadget of color combination $i,j$} with $i<j$ has the following sets.
    \begin{itemize}
        \item Sets $A$ and $B$ contain all vertices in $X_{i,j}$ and vertex $v_{i,j}^{\star\star}$.
        %\item Set $A'$ contains all vertices from all edge selection gadgets and all validation gadgets except $X_{i,j}$.
        %\item Set $B$ contains all vertices from all edge selection gadgets and all validation gadgets, except $v_{i,j}^{\star\star}$.
        %\item Set $B$ contains vertex $v_{i,j}^{\star\star}$.
    \end{itemize}
    The \emph{second connector gadget for the edge selection gadget of color combination $i,j$} with $i<j$ has the following sets.
    \begin{itemize}
        \item Set $A$ contains all vertices from the edge selection gadget $G_{i,j}$ except vertices in $X_{i,j}$.
        %\item Set $A'$ contains all vertices in $X_{i,j}$.
        %\item Set $B$ contains all vertices from all edge selection gadgets and all validation gadgets.
        \item Set $B$ is empty.
    \end{itemize}
    The \emph{first connector gadget for the verification gadget of color $i$} has the following sets.
    \begin{itemize}
        %\item Set $A$ contains all vertices of the verification gadget. % and all vertices from edge selection gadgets that are neighbor of a vertex in the verification gadget.
        \item Sets $A$ and $B$ contain all vertices $v^i_\ell$ with $0\le \ell\le k$.
    \end{itemize}
The \emph{second connector gadget for the verification gadget of color $i$} has the following sets.
    \begin{itemize}
        \item Set $A$ contains all vertices of the verification gadget except vertices $v^i_\ell$ with $0\le \ell \le k$.
        \item Set $B$ is empty.
    \end{itemize}

\emph{Alignment Gadget.} Lastly, we introduce an \emph{alignment gadget}. It consists of one vertex $w^\star$ and a set of vertices $\hat{W}$ containing one vertex for each edge selection gadget, one vertex for each verification gadget, and one vertex for each connector gadget. Vertex $w^\star$ is connected to each vertex in $\hat{W}$.
The vertex $x_1$ of each edge selection gadget, the vertex $y^i$ of each verification gadget, and the vertex $\hat{v}_1$ of each connector gadget are each connected to  one vertex in $\hat{W}$ such that all vertices in $\hat{W}$ have degree two.

\begin{figure}[t!]
\noindent\makebox[\textwidth]{
\centering
%\begin{center}
\begin{tikzpicture}[line width=1pt, scale=.65, yscale=.7]
\scriptsize
\node (A) at (14,0) {};

\draw[rounded corners, dashed, gray] (-13, 5.4) rectangle (10, 1.6) {};
\node (A) at (-10.5,4.5) {alignment gadget};

\draw[rounded corners, dotted, gray] (-14, -2.5) rectangle (-10.5, -9.5) {};
\draw[rounded corners, dotted, gray] (-14, -12.5) rectangle (-10.5, -19.5) {};
\draw[rounded corners, dotted, gray] (7.5, -2.5) rectangle (11, -9.5) {};
\draw[rounded corners, dotted, gray] (7.5, -12.5) rectangle (11, -19.5) {};
\node[rotate=90] (A) at (-14.5,-16) {connector gadget};

\draw[rounded corners, dashed, gray] (-9.5, 1.4) rectangle (1.5, -26) {};
\node (A) at (-4,-25.5) {edge selection gadget for $i,j$};

\draw[rounded corners, dashed, gray] (3.5, 1.4) rectangle (6.5, -26) {};
\node[rotate=-90] (A) at (7,-21.5) {verification gadget for $j$};

    \node[vert,label=above:$w^\star$, fill=gray] (W) at (-2,4) {};
    \node[vert] (W1) at (-6.5,2) {};
    \node[vert] (W2) at (-10,2) {};
    \node[vert] (W3) at (-12,2) {};
    \node[vert] (W4) at (7,2) {}; 
    \node[vert] (W5) at (3.5,2) {};
    \node[vert] (W6) at (9,2) {};

    \node[vert,label=left:$\hat{v}_0$] (VH11) at (-11,-3) {}; 
    \node[vert,label=left:$\hat{v}_0'$] (VH111) at (-13,-3) {}; 
    \node[vert,label=left:$\hat{v}_1$, fill=gray] (VH12) at (-12,-5) {}; 
    \node[vert,label=left:$\hat{v}_2$, fill=gray] (VH13) at (-12,-7) {}; 
    \node[vert,label=left:$\hat{v}_3$] (VH14) at (-11,-9) {};  
    \node[vert,label=left:$\hat{v}_3'$] (VH144) at (-13,-9) {};

    \node[vert,label=left:$\hat{v}_0$] (VH21) at (-11,-13) {};
    \node[vert,label=left:$\hat{v}_0'$] (VH211) at (-13,-13) {}; 
    \node[vert,label=left:$\hat{v}_1$, fill=gray] (VH22) at (-12,-15) {}; 
    \node[vert,label=left:$\hat{v}_2$, fill=gray] (VH23) at (-12,-17) {};
    \node[vert,label=left:$\hat{v}_3$] (VH24) at (-11,-19) {};
    \node[vert,label=left:$\hat{v}_3'$] (VH244) at (-13,-19) {};

    \node[vert,label=right:$\hat{v}_0$] (VH31) at (8,-3) {}; 
    \node[vert,label=right:$\hat{v}_0'$] (VH311) at (10,-3) {}; 
    \node[vert,label=right:$\hat{v}_1$, fill=gray] (VH32) at (9,-5) {}; 
    \node[vert,label=right:$\hat{v}_2$, fill=gray] (VH33) at (9,-7) {}; 
    \node[vert,label=right:$\hat{v}_3$] (VH34) at (8,-9) {};  
    \node[vert,label=right:$\hat{v}_3'$] (VH344) at (10,-9) {};

    \node[vert,label=right:$\hat{v}_0$] (VH41) at (8,-13) {}; 
    \node[vert,label=right:$\hat{v}_0'$] (VH411) at (10,-13) {}; 
    \node[vert,label=right:$\hat{v}_1$, fill=gray] (VH42) at (9,-15) {}; 
    \node[vert,label=right:$\hat{v}_2$, fill=gray] (VH43) at (9,-17) {}; 
    \node[vert,label=right:$\hat{v}_3$] (VH44) at (8,-19) {};  
    \node[vert,label=right:$\hat{v}_3'$] (VH444) at (10,-19) {};

    \node[vert,label=above:$x_1$] (X1) at (-8,0) {}; 
    \node[vert,label=above:$x_2$] (X2) at (-6,0) {}; 
    \node[vert,label=above:$x_3$] (X3) at (-4,0) {}; 
    \node (A) at (-2,0) {$\ldots$};
    \node[vert,label=above:$x_m$] (XM) at (0,0) {};

    \node[vert,label=left:$v_{i,j}^\star$, fill=gray] (VS) at (-4,-3) {};

    \node[vert,label=left:$u^1_0$] (Y1) at (-8,-6) {}; 
    \node[vert,label=left:$u^2_0$] (Y2) at (-6,-6) {}; 
    \node[vert,label=left:$u^3_0$] (Y3) at (-4,-6) {}; 
    \node (A) at (-2,-6) {$\ldots$};
    \node[vert,label=right:$u^m_0$] (YM) at (0,-6) {};

    \node[vert,label=left:$u^1_1$] (U11) at (-8,-8) {}; 
    \node[vert,label=left:$u^2_1$] (U12) at (-6,-8) {}; 
    \node[vert,label=left:$u^3_1$] (U13) at (-4,-8) {}; 
    \node (A) at (-2,-8) {$\ldots$};
    \node[vert,label=right:$u^m_1$] (U1M) at (0,-8) {};

    \node[vert,label=left:$u^1_2$] (U21) at (-8,-10) {}; 
    \node[vert,label=left:$u^2_2$] (U22) at (-6,-10) {}; 
    \node[vert,label=left:$u^3_2$] (U23) at (-4,-10) {}; 
    \node (A) at (-2,-10) {$\ldots$};
    \node[vert,label=right:$u^m_2$] (U2M) at (0,-10) {};

    \node (A) at (-8,-11.5) {$\vdots$}; 
    \node (A) at (-6,-11.5) {$\vdots$}; 
    \node (A) at (-4,-11.5) {$\vdots$}; 
    \node (A) at (0,-11.5) {$\vdots$};

    \node[vert,label=left:$u^1_n$] (UN1) at (-8,-13) {}; 
    \node[vert,label=left:$u^2_n$] (UN2) at (-6,-13) {}; 
    \node[vert,label=left:$u^3_n$] (UN3) at (-4,-13) {}; 
    \node (A) at (-2,-13) {$\ldots$};
    \node[vert,label=right:$u^m_n$] (UNM) at (0,-13) {};

    \node[vert,label=left:$u_{n+1}^1$] (UN11) at (-8,-17) {}; 
    \node[vert,label=left:$u_{n+1}^2$] (UN12) at (-6,-17) {}; 
    \node[vert,label=left:$u_{n+1}^3$] (UN13) at (-4,-17) {}; 
    \node (A) at (-2,-17) {$\ldots$};
    \node[vert,label=right:$u_{n+1}^m$] (UN1M) at (0,-17) {};

    \node (A) at (-8,-18.5) {$\vdots$}; 
    \node (A) at (-6,-18.5) {$\vdots$}; 
    \node (A) at (-4,-18.5) {$\vdots$}; 
    \node (A) at (0,-18.5) {$\vdots$};

    \node[vert,label=left:$u_{4n}^1$] (U2N1) at (-8,-20) {}; 
    \node[vert,label=left:$u_{4n}^2$] (U2N2) at (-6,-20) {}; 
    \node[vert,label=left:$u_{4n}^3$] (U2N3) at (-4,-20) {}; 
    \node (A) at (-2,-20) {$\ldots$};
    \node[vert,label=right:$u_{4n}^m$] (U2NM) at (0,-20) {};

    \node[vert,label=left:$v_{i,j}^{\star\star}$, fill=gray] (VSS) at (-4,-24) {};

    \node[vert,label=right:$y^j$] (YI) at (5,0) {};
    \node[vert,label=right:$v_0^j$, fill=gray] (VI0) at (5,-2) {};
    \node[vert,label=right:$v_1^j$, fill=gray] (VI1) at (5,-4) {};
    \node (A) at (5,-5.5) {$\vdots$};
    \node[vert,label=right:$v_{i-1}^j$, fill=gray] (VIJ1) at (5,-7) {};
    \node[vert,label=right:$v_{i}^j$, fill=gray] (VIJ) at (5,-11) {};
    \node (A) at (5,-12.5) {$\vdots$};
    \node[vert,label=right:$v_{j-1}^j$, fill=gray] (VII1) at (5,-14) {};
    \node[vert,label=right:$\hat{u}_{1}^j$] (UI1) at (5,-16) {};
    \node[vert,label=right:$\hat{u}_{2}^j$] (UI2) at (5,-18) {};
    \node (A) at (5,-19.5) {$\vdots$};
    \node[vert,label=right:$v_{j}^j$, fill=gray] (UIN3) at (5,-21) {};
    \node (A) at (5,-22.5) {$\vdots$};
    \node[vert,label=right:$v_{k}^j$, fill=gray] (VIK) at (5,-24) {};

	\draw[edge] (X1) --node[] {} (VS);
	\draw[edge] (X2) --node[] {} (VS);
	\draw[edge] (X3) --node[] {} (VS);
	\draw[edge] (XM) --node[] {} (VS);

    \draw[edge] (Y1) --node[] {} (VS);
	\draw[edge] (Y2) --node[] {} (VS);
	\draw[edge] (Y3) --node[] {} (VS);
	\draw[edge] (YM) --node[] {} (VS);

    \draw[edge] (Y1) --node[] {} (U11);
	\draw[edge] (Y2) --node[] {} (U12);
	\draw[edge] (Y3) --node[] {} (U13);
	\draw[edge] (YM) --node[] {} (U1M);

    \draw[edge] (U21) --node[] {} (U11);
	\draw[edge] (U22) --node[] {} (U12);
	\draw[edge] (U23) --node[] {} (U13);
	\draw[edge] (U2M) --node[] {} (U1M);

    \draw[edge] (UN1) --node[] {} (UN11);
	\draw[edge] (UN2) --node[] {} (UN12);
	\draw[edge] (UN3) --node[] {} (UN13);
	\draw[edge] (UNM) --node[] {} (UN1M);

    \draw[edge] (U2N1) --node[] {} (VSS);
	\draw[edge] (U2N2) --node[] {} (VSS);
	\draw[edge] (U2N3) --node[] {} (VSS);
	\draw[edge] (U2NM) --node[] {} (VSS);

    \draw[edge] (VH11) --node[] {} (VH12);
	\draw[edge] (VH12) --node[] {} (VH111);
	\draw[edge] (VH12) --node[] {} (VH13);
    \draw[edge] (VH13) --node[] {} (VH14);
	\draw[edge] (VH13) --node[] {} (VH144);

    \draw[edge] (VH21) --node[] {} (VH22);
	\draw[edge] (VH22) --node[] {} (VH211);
	\draw[edge] (VH22) --node[] {} (VH23);
    \draw[edge] (VH23) --node[] {} (VH24);
	\draw[edge] (VH23) --node[] {} (VH244);

    \draw[edge] (VH31) --node[] {} (VH32);
	\draw[edge] (VH32) --node[] {} (VH311);
	\draw[edge] (VH32) --node[] {} (VH33);
    \draw[edge] (VH33) --node[] {} (VH34);	
    \draw[edge] (VH33) --node[] {} (VH344);

    \draw[edge] (VH41) --node[] {} (VH42);
	\draw[edge] (VH42) --node[] {} (VH411);
	\draw[edge] (VH42) --node[] {} (VH43);
    \draw[edge] (VH43) --node[] {} (VH44);	
    \draw[edge] (VH43) --node[] {} (VH444);

	%\draw[edge] (VIJ1) --node[timelabel, near start] {$t_1+4$} (U13);
	%\draw[edge] (VIJ) --node[timelabel, near start] {$t_1+8$} (U23);

    \draw[edge,double] (VIJ1) --node {} (U13);
	\draw[edge,double] (VIJ) --node {} (U23);

	\draw[edge] (W) --node[] {} (W1);
	\draw[edge] (W) --node[] {} (W2);
	\draw[edge] (W) --node[] {} (W3);
	\draw[edge] (W) --node[] {} (W4);
	\draw[edge] (W) --node[] {} (W5);
	\draw[edge] (W) --node[] {} (W6);

	\draw[edge] (X1) --node[] {} (W1);
	\draw[edge] (VH11) --node[] {} (W2);
	\draw[edge] (VH21) edge[bend left=50] node[] {} (W3);
	\draw[edge] (VH31) --node[] {} (W4);
	\draw[edge] (YI) --node[] {} (W5);
	\draw[edge] (VH41) edge[bend right=50] node[] {} (W6);

	\draw[edge] (YI) --node[] {} (VI0);

	\draw[edge] (VII1) --node[] {} (UI1);
	\draw[edge] (UI1) --node[] {} (UI2);
\end{tikzpicture}
    %\end{center}
    }
    \caption{Illustration of part of the underlying graph $G$. Edges incident with vertices $\hat{v}_1,\hat{v}_2$ of connector gadgets are omitted. Gray vertices form a feedback vertex set.
    The double line connections, between a vertex $v_{i-1}^j$ in the verification gadget, and $u_1^3$ in the edge selection gadget, 
    and, between a vertex $u_2^3$ in the edge selection gadget, and $v_{i}^j$ in the verification gadget,
    consist of $5n$ vertices $a_1^{j,i,3},a_2^{j,i,3},\dots,a_{5n}^{j,i,3}$ 
    and $b_1^{j,i,3},b_2^{j,i,3},\dots,b_{5n}^{j,i,3}$, respectively.
    }\label{fig:hardness1}
\end{figure}

@@

This finishes the description of the underlying graph $G$. For an illustration see \cref{fig:hardness1}. We can observe that the vertex set containing
%\begin{itemize}
 %   \item 
    vertices $v_{i,j}^\star$ and $v_{i,j}^{\star\star}$ of each edge selection gadget,
    %\item 
    vertices $v^i_\ell$ with $0\le \ell\le k$ of each verification gadget,
    %\item 
    vertices $\hat{v}_1$ and $\hat{v}_2$ of each connector gadget, and
    %\item 
    vertex $w^\star$ of the alignment gadget
%\end{itemize}
forms a feedback vertex set in $G$ with size $O(k^2)$.

As mentioned before, due to space constraints, we defer the description of the distance matrix $D$ and a formal correctness proof of the reduction to the Appendix.

\begin{comment}
\emph{Duration Matrix $D$.} We proceed with describing the matrix $D$ of durations of fastest paths. For a more convenient presentation, we use the notation $d(v,v'):= D_{v,v'}$. For all vertices $v,v'$ that are neighbors in $G$ we have that $d(v,v')=1$ and $d(v',v)=1$.

Next, consider a connector gadget consisting of vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$ and with sets $A$ and $B$. Informally, the connector gadget makes sure that all vertices in $A$ can reach all other vertices (of edge selection gadgets and validation gadgets) except the ones in $B$. We set the following durations. Recall that $V^\star$ denotes the set of all vertices from all edge selection gadgets and all validation gadgets.
\begin{itemize}
    \item We set $d(\hat{v}_0,\hat{v}_2)=d(\hat{v}_3,\hat{v}_1)=d(\hat{v}_2,\hat{v}_0')=d(\hat{v}_1,\hat{v}_3')=2$, and $d(\hat{v}_0,\hat{v}_0')=d(\hat{v}_3,\hat{v}_3')=d(\hat{v}_0,\hat{v}_3')=d(\hat{v}_3,\hat{v}_0')=3$.
    \item Let $v\in A$, then we set $d(v,\hat{v}_0')=3$ and $d(v,\hat{v}_3')=3$.
    \item Let $v\in V^\star\setminus B$, then we set $d(\hat{v}_0,v)=3$ and $d(\hat{v}_3,v)=3$.
    \item Let $v\in A$ and $v'\in V^\star\setminus B$ such that $v$ and $v'$ are not neighbors, then we set $d(v,v')=3$.
\end{itemize}
Now consider two connector gadgets, one with vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$ and with sets $A$ and $B$, and one with vertices $\hat{v}_0',\hat{v}_0'',\hat{v}_1',\hat{v}_2',\hat{v}_3',\hat{v}_3''$ and with sets $A'$ and $B'$.
\begin{itemize}
    \item If there is a vertex $v\in A$ with $v\notin A'$, then we set $d(\hat{v}_1,\hat{v}_1')=3$.
    \item If there is a vertex $v\in A$ with $v\in A'\setminus B'$, then we set $d(\hat{v}_1,\hat{v}_2')=3$.
    \item If there is a vertex $v\in V^\star\setminus (A\setminus B)$ with $v\notin A'$, then we set $d(\hat{v}_2,\hat{v}_1')=3$.    
    \item If there is a vertex $v\in V^\star\setminus (A\setminus B)$ with $v\in A'\setminus B'$, then we set $d(\hat{v}_2,\hat{v}_2')=3$.
\end{itemize}

Next, consider the edge selection gadget for color combination $i,j$ with $i<j$.
\begin{itemize}
    \item Let $1\le \ell<\ell'\le m$. We set $d(x_\ell,x_{\ell'})=2n\cdot (i+j)\cdot((\ell')^2-\ell^2)+1$.
    \item For all $\ell\in[m]$ we set $d(x_\ell,v_{i,j}^{\star\star})=8n+5$.
\end{itemize}

Next, consider the validation gadget for color $i$. 
%Let $1\le j<i$, let $e_\ell\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. We set the following.
%\begin{itemize}
    %\item We set $d(v^i_{j-1},w^i_{\ell'+1})=3$.
    %\item We set $d(w^i_{\ell'},v^i_{j})=3$.
%\end{itemize}
%
For all $0\le j<j'<i$ and all $i\le j<j'\le k$ we set the following.
\begin{itemize}
    \item We set $d(v^i_j,v^i_{j'})=(20n+6)(j'-j)-1$.
\end{itemize}
%The next cases are only relevant if $1\neq i \neq k$. 
%For all $0\le j<i$ %and $\ell\in[n+3]$ we set the following.
%\begin{itemize}
%    \item We set $d(v^i_j,\hat{u}^i_{n+3})=(20n+6)(i-j-1)+2n+5$.
%\end{itemize}
%For all $\ell,\ell'\in[n+3]$ with $\ell<\ell'$ we set the following.
%\begin{itemize}
%    \item We set $d(\hat{u}^i_{\ell},\hat{u}^i_{\ell'})=2(\ell'-\ell)-1$.
%\end{itemize}
For all $0\le j<i$ and all $i\le j'\le k$ we set the following.
\begin{itemize}
    \item We set $d(v^i_j,v^i_{j'})=(20n+6)(j'-j)+6n-1$.
\end{itemize}
%For all 
%$\ell\in[n+3]$ and 
%$i<j\le k$ we set the following.
%\begin{itemize}
%    \item We set $d(\hat{u}^i_{n+3},v^i_j)=(20n+6)(j-i)-1$.
%\end{itemize}

Finally, we consider the alignment gadget. Let $x_1$ belong to the edge selection gadget of color combination $i,j$ and let $w\in \hat{W}$ denote the neighbor of $x_1$ in the alignment gadget. Let $\hat{v}_1$ and $\hat{v}_2$ belong to the first connector gadget of the edge selection gadget for color combination $i,j$. Let $\hat{V}$ contain all vertices $\hat{v}_1$ and $\hat{v}_2$ belonging to the other connector gadgets (different from the first one of the edge selection gadget for color combination $i,j$). 
\begin{itemize}
    \item We set $d(w^\star,x_1)=(20n+6)(i+j)$.
    \item We set $d(w^\star,\hat{v}_1)=n^9$, $d(w,\hat{v}_2)=n^9$, $d(w,\hat{v}_1)=n^9-(20n+6)(i+j)+1$, and $d(w,\hat{v}_2)=n^9-(20n+6)(i+j)+1$.
    \item For each vertex $v\in (V^\star\cup \hat{V})\setminus (X_{i,j}\cup \{v_{i,j}^{\star\star}\})$ we set $d(w^\star,v)=n^9+2$ and $d(w,v)=n^9-(20n+6)(i+j)+3$.
\end{itemize}

Let $y^i$ belong to the verification gadget of color $i$ and let $w'\in \hat{W}$ denote the neighbor of $y^i$ in the alignment gadget. Let $\hat{v}_1$ and $\hat{v}_2$ belong to the connector gadget of the verification gadget for color $i$. Let $\hat{V}$ contain all vertices $\hat{v}_1$ and $\hat{v}_2$ belonging to the other connector gadgets (different from the one of the verification gadget for color $i$). Let $V_i$ denote the set of all vertices of the verification gadget of color $i$.
\begin{itemize}
    \item We set $d(w^\star,y^i)=n^8-1$, $d(w',v^i_0)=2$, and $d(w^\star,v^i_0)=n^8$.
    \item We set $d(w^\star,\hat{v}_1)=n^9$, $d(w^\star,\hat{v}_2)=n^9$, $d(w',\hat{v}_1)=n^9-n^8$, and $d(w',\hat{v}_2)=n^9-n^8$.
    \item For each vertex $v\in (V^\star\cup \hat{V})\setminus V_i$ we set $d(w^\star,v)=n^9+1$, $d(w,v)=n^9-n^8+2$, and $d(y^i,v)=n^9-n^8+2$.
\end{itemize}
Let $\hat{v}_1$ belong to some connector gadget. Then we set $d(w^\star,\hat{v}_1)=n^9$.

All fastest path durations between non-adjacent vertex pairs that are not specified above are set to infinity.

%This finishes the construction of \deltaExactLong\ instance $D$, which can clearly be computed in polynomial time. For an illustration see \cref{fig:hardness1}. As discussed earlier, we have that the vertex cover number of the underlying graph of the instance is in $O(k^2)$.

>>>>

@@

We obtain the following results. We show that \deltaExact\ is W[1]-hard when parameterized by the \emph{feedback vertex number} of the underlying graph (\cref{thm:W1wrtFVS}). 
To this end, we first give a reduction from \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval} to a variant of \deltaExact\ where the period $\Delta$ is infinite, that is, when the labeling is non-periodic. 
Then we use a special gadget (the ``infinity'' gadget) which allows us to transfer the result to a finite period $\Delta$. The latter construction is independent from the particular reduction we use, and can hence be treated as a reduction from the non-periodic to the periodic setting.
Note that our parameterized hardness result with respect to the feedback vertex number 
also implies W[1]-hardness for any smaller parameter, such as \emph{treewidth}, \emph{degeneracy}, \emph{cliquewidth}, 
\emph{distance to chordal graphs}, and \emph{distance to outerplanar graphs}.

@@

%There are some problem settings that share similarities with ours, which we discuss now in more detail.

@@

\begin{theorem}[$\star$]\label{thm:W1wrtFVS}
    \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$ (if not we can add $k \max_{i,j}|F_{i,j}|$ vertices to each $W_i$ and use those to add up to $\max_{i,j}|F_{i,j}|$ additional isolated edges to each $F_{i,j}$).
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

We give a reduction to a variant of \deltaExact\ where the period $\Delta$ is infinite (that is, the sought temporal graph is not periodic and the labeling function $\lambda : E\rightarrow \mathbb{N}$ maps to the natural numbers) and we allow $D$ to have infinity entries, meaning that the two respective vertices are not temporally connected. 
Note that, given the matrix $D$, we can easily compute the underlying graph $G$, as follows. Two vertices $v,v'$ are adjacent if $G$ if and only if $D_{v,v'}=1$, as having an edge between $v$ and $v'$ is the only way that there exists a temporal path from $v$ to $v'$ with duration 1. 
For simplicity of the presentation of the reduction, we describe the underlying graph $G$ (which directly implies the entries of $D$ where $D(v,v')=1$) and then we provide the remaining entries of $D$. 
In the Appendix, we show how to obtain the result for a finite $\Delta$ (by introducing a so-called ``infinity gadget'') and a matrix $D$ of durations of fastest paths which only has finite entries.

@@

Every \emph{edge selection gadget} is associated with a color combination $i,j$ in the \textsc{Multicolored Clique} instance, and its main purpose is to ``select'' an edge connecting a vertex from color $i$ with a vertex from color $j$.
Roughly speaking, the edge selection gadget consists of $m$ paths, one for every edge in $F_{i,j}$ (see Figure~3 in the Appendix for reference). The distance matrix $D$ will enforce that the labels on those paths effectively order them temporally, that is, in particular, the labels on one of the paths will be smaller than the labels on all other paths. The edge corresponding to this path is selected.

 We have a \emph{verification gadget} for every color $i$. They interact with the edge selection gadgets as follows.
 The verification gadget for color $i$ is connected to all edge selection gadgets that involve color $i$. More specifically, this is connected to every path corresponding to an edge at a position in the path that encodes the endpoint of color $i$ of that edge (again, see Figure~3 in the Appendix for reference). Intuitively, the distances in the verification gadget are only realizable if the selected edges all have the same endpoint of color $i$.
Hence, the distances of all verification gadgets can be realized if and only if the selected edges form a clique.

@@

@@

This finishes the description of the underlying graph $G$. For an illustration see Figure~3 in the Appendix. We can observe that the vertex set containing
%\begin{itemize}
 %   \item 
    vertices $v_{i,j}^\star$ and $v_{i,j}^{\star\star}$ of each edge selection gadget,
    %\item 
    vertices $v^i_\ell$ with $0\le \ell\le k$ of each verification gadget,
    %\item 
    vertices $\hat{v}_1$ and $\hat{v}_2$ of each connector gadget, and
    %\item 
    vertex $w^\star$ of the alignment gadget
%\end{itemize}
forms a feedback vertex set in $G$ with size $O(k^2)$.

As mentioned before, due to space constraints, we defer the description of the distance matrix $D$ and a formal correctness proof of the reduction to the Appendix.
\end{proof}
